Online Multiclass Linear Classification with Bandit Feedback

### 2018-11-28 arxiv-version/main.tex
Finish proof of properties of Chebyshev polynomials

However, for $K \ge 3$ there exists a set of labeled examples in an inner product
space that is linearly separable with a positive margin $\gamma$ but it is not
strongly linearly separable with any positive margin $\gamma$. Such set of labeled
examples is shown in
Figure~\ref{figure:linearly-separable-examples-with-margin}.

@@

The idea behind the algorithm is to use $K$ copies of the \textsc{Binary
Perceptron} algorithm. Each copy corresponds to a class. In each round, each
copy makes a binary decision whether the feature vector belongs their respective
class or not. If at least one copy makes a positive prediction, the algorithm
predicts accordingly. If multiple copies make positive predictions, the
algorithm chooses one of them at random. If all copies make negative
predictions, the algorithms makes a blind guess.

@@

\begin{theorem}[Mistake upper bound]
\label{theorem:strongly-separable-example-mistake-upper-bound}
Let $(V, \ip{\cdot}{\cdot})$ be any inner product space, let $K$ be a positive
integer, $\gamma$ a positive real number, $R$ be a non-negative real number. If
$(x_1, y_1), (x_2, y_2), \dots, (x_T, y_T)$ is a sequence of labeled examples in
$V \times \{1,2,\dots,K\}$ such that the examples are strongly separable with
margin $\gamma$ and $\norm{x_1}, \norm{x_2}, \dots, \norm{x_T} \le R$ then the
expected number of mistakes
Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples}
makes is at most $(K-1) \lfloor (R/\gamma)^2 \rfloor$ mistakes.
\end{theorem}

@@

\begin{theorem}[Mistake lower bound]
\label{theorem:strongly-separable-example-mistake-lower-bound}
Let $\gamma$ be a positive real number and let $R$ be a non-negative real
number. For any algorithm for the \textsc{Online Multiclass Classification with
Bandit Feedback} with $K \le \frac{1}{4}(R/\gamma)^2$ classes there exists an inner
product space $(V, \ip{\cdot}{\cdot})$, a non-negative integer $T$ and a
sequence of labeled examples $(x_1, y_1), (x_2, y_2), \dots, (x_T, y_T)$
in $V \times \{1,2,\dots,K\}$ that are strongly linearly separable with
margin $\gamma$, the norms satisfy $\norm{x_1}, \norm{x_2}, \dots, \norm{x_T}
\le R$ and the expected number of mistakes the algorithm makes is at least $\frac{K-1}{2}
\lfloor \frac{1}{16} (R/\gamma)^2 \rfloor$.
\end{theorem}

@@

Finally, since strong separability and norm condition hold with probability one,
there exists a particular (i.e. deterministic) sequence of examples for which
the algorithm makes at least $\frac{K-1}{2} M$ mistakes.
\end{proof}

@@

The idea behind all the mappings is that, on a compact set, multivariate
polynomials uniformly approximate any continuous function. This result is a
special case of the well known Stone-Weierstrass theorem; see
e.g.~\citep[Section~10.10]{Davidson-Donsig-2010}. In our case, we use
multivariate polynomials to approximate the indicator function corresponding to
the intersection of $K-1$ halfspaces. This polynomial separates examples in one
class from examples in the other classes. To be able to quantify $\gamma'$ and
$R'$ as well as time complexity of evaluating $k(x,x')$ we need to quantify
certain parameters of the approximating polynomial (degree, norm, range of
values). The parameters are quantified in
\autoref{theorem:polynomial-approximation-1} and
\autoref{theorem:polynomial-approximation-2} stated below.

Before we state the theorem, recall that a polynomial of $d$ variables is a
function $p:\R^d \to \R$ of the form
$$
p(x) = p(x_1, x_2, \dots, x_d) = \sum_{i_1, i_2, \dots, i_d} c_{i_1, i_2, \dots, i_d} x_1^{i_1} x_2^{i_2} \dots x_d^{i_d}
$$
where the sum ranges over a finite set of tuples $(i_1, i_2, \dots, i_d)$ of
non-negative integers and $c_{i_1, i_2, \dots, i_d}$ is a real coefficient. The
\emph{degree} of a polynomial $p$, denoted by $\deg(p)$, is the largest value of
$i_1 + i_2 + \dots + i_d$ for which the coefficient is non-zero. The \emph{norm
of a polynomial} $p$ is defined as
$$
\norm{p} = \sqrt{\sum_{i_1, i_2, \dots, i_d} \left(c_{i_1, i_2, \dots, i_d} \right)^2 } \; .
$$

@@

Note that $\cosh(\theta) = \frac{e^{\theta} + e^{-\theta}}{2}$ is an even
continuous function that maps $\R$ onto $[1,+\infty)$, is strictly decreasing
on $(-\infty,0]$, and is strictly increasing on $[0,\infty)$. Thus the fourth
property is equivalent to
$$
T_n(\cosh(\theta)) \ge 1 + n^2 (\cosh(\theta) - 1) \qquad \text{for any $\theta \ge 0$.}
$$
For $\theta = 0$, according to \eqref{equation:hyperbolic-chebyshev} both sides
are equal to $1$. The proof will be finished, if we prove that the derivative of
the left hand side is greater or equal to the derivative of the right hand side.
That is, we need to prove that
\begin{equation}
\label{equation:hyperbolic-chebyshev-derivative-inequality}
\frac{\diff T_n(\cosh(\theta))}{\diff \theta} \ge \frac{\diff (1 + n^2 (\cosh(\theta) - 1))}{\diff \theta} \qquad \text{for any $\theta \ge 0$.}
\end{equation}
Using \eqref{equation:hyperbolic-chebyshev} and recalling that $\diff \cosh(\theta) / \diff \theta = \sinh(\theta)$ we
can express the derivatives as
\begin{align*}
\frac{\diff T_n(\cosh(\theta))}{\diff \theta} & = \frac{\diff \cosh(n \theta))}{\diff \theta} = n \sinh(n \theta) \; , \\
\frac{\diff (1 + n^2 (\cosh(\theta) - 1))}{\diff \theta} & = n^2 \sinh(\theta) \; .
\end{align*}
TODO

>>>>

However, for any $K \ge 3$ and any inner product space of dimension at least
$2$, there exists a set of labeled examples that is linearly separable with a
positive margin $\gamma$ but it is not strongly linearly separable with any
positive margin $\gamma$. An example of such set of labeled examples is shown in
Figure~\ref{figure:linearly-separable-examples-with-margin}.

@@

The idea behind the algorithm is to use $K$ copies of the \textsc{Binary
Perceptron} algorithm. Each copy corresponds to a class. In each round, each
copy makes a binary decision whether the feature vector belongs its respective
class or not. If at least one copy makes a positive prediction, the algorithm
predicts accordingly. If multiple copies make positive predictions, the
algorithm chooses one of them arbitrarily. If all copies make negative
predictions, the algorithms makes a blind guess.

@@

\begin{theorem}[Mistake upper bound]
\label{theorem:strongly-separable-example-mistake-upper-bound}
Let $(V, \ip{\cdot}{\cdot})$ be any inner product space, let $K$ be a positive
integer, $\gamma$ a positive real number, $R$ be a non-negative real number. If
$(x_1, y_1), (x_2, y_2), \dots, (x_T, y_T)$ is a sequence of labeled examples in
$V \times \{1,2,\dots,K\}$ such that the examples are strongly separable with
margin $\gamma$ and $\norm{x_1}, \norm{x_2}, \dots, \norm{x_T} \le R$ then the
expected number of mistakes
Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples}
makes is at most $(K-1) \lfloor (R/\gamma)^2 \rfloor$.
\end{theorem}

@@

\begin{theorem}[Mistake lower bound]
\label{theorem:strongly-separable-example-mistake-lower-bound}
Let $\gamma$ be a positive real number and let $R$ be a non-negative real
number. For any (possibly randomized) algorithm for the \textsc{Online
Multiclass Classification with Bandit Feedback} with $K \le
\frac{1}{4}(R/\gamma)^2$ classes there exists an inner product space $(V,
\ip{\cdot}{\cdot})$, a non-negative integer $T$ and a sequence of labeled
examples $(x_1, y_1), (x_2, y_2), \dots, (x_T, y_T)$ in $V \times
\{1,2,\dots,K\}$ that are strongly linearly separable with margin $\gamma$, the
norms satisfy $\norm{x_1}, \norm{x_2}, \dots, \norm{x_T} \le R$ and the expected
number of mistakes the algorithm makes is at least $\frac{K-1}{2} \lfloor
\frac{1}{16} (R/\gamma)^2 \rfloor$.
\end{theorem}

@@

Finally, since strong separability and norm condition hold with probability one,
there exists a particular (i.e. deterministic) sequence of examples for which
the algorithm makes at least $\frac{K-1}{2} M$ mistakes in expectation
over its internal randomization.
\end{proof}

@@

The idea behind all the mappings is that, on a compact set, multivariate
polynomials uniformly approximate any continuous function. This result is a
special case of the well known Stone-Weierstrass theorem; see
e.g.~\citep[Section~10.10]{Davidson-Donsig-2010}. In our case, we use
multivariate polynomial to approximate the indicator function corresponding to
the intersection of $K-1$ halfspaces. This polynomial separates examples in one
class from examples in the other classes. To be able to quantify $\gamma'$ and
$R'$ as well as time complexity of evaluating $k(x,x')$ we need to quantify
certain parameters of the approximating polynomial (degree, norm, range of
values). We construct two different polynomials with different paremeters. The
parameters are quantified in
Theorems~\ref{theorem:polynomial-approximation-1}~and~\ref{theorem:polynomial-approximation-2}
stated below.

Before we state the theorems, recall that a polynomial of $d$ variables is a
function $p:\R^d \to \R$ of the form
$$
p(x) = p(x_1, x_2, \dots, x_d) = \sum_{i_1, i_2, \dots, i_d} c_{i_1, i_2, \dots, i_d} x_1^{i_1} x_2^{i_2} \dots x_d^{i_d}
$$
where the sum ranges over a finite set of tuples $(i_1, i_2, \dots, i_d)$ of
non-negative integers and $c_{i_1, i_2, \dots, i_d}$ is a real coefficient. The
\emph{degree} of a polynomial $p$, denoted by $\deg(p)$, is the largest value of
$i_1 + i_2 + \dots + i_d$ for which the coefficient is non-zero. The \emph{norm
of a polynomial} $p$ is defined as
$$
\norm{p} = \sqrt{\sum_{i_1, i_2, \dots, i_d} \left(c_{i_1, i_2, \dots, i_d} \right)^2 } \; .
$$

@@

Note that $\cosh(\theta) = \frac{e^{\theta} + e^{-\theta}}{2}$ is an even
continuous function that maps $\R$ onto $[1,+\infty)$, is strictly decreasing
on $(-\infty,0]$, and is strictly increasing on $[0,\infty)$. Thus the fourth
property is equivalent to
$$
T_n(\cosh(\theta)) \ge 1 + n^2 (\cosh(\theta) - 1) \qquad \text{for all $\theta \ge 0$.}
$$
For $\theta = 0$, according to \eqref{equation:hyperbolic-chebyshev} both sides
are equal to $1$. Thus, it is sufficient to prove that the derivative of the
left hand side is greater or equal to the derivative of the right hand side.
That is, we need to prove that
\begin{equation}
\label{equation:hyperbolic-chebyshev-derivative-inequality}
\frac{\diff T_n(\cosh(\theta))}{\diff \theta} \ge \frac{\diff (1 + n^2 (\cosh(\theta) - 1))}{\diff \theta} \qquad \text{for all $\theta \ge 0$.}
\end{equation}
Using \eqref{equation:hyperbolic-chebyshev} and recalling that $\frac{\diff \cosh(\theta)}{\diff \theta} = \sinh(\theta)$ we
can express the derivatives as
\begin{align*}
\frac{\diff T_n(\cosh(\theta))}{\diff \theta} & = \frac{\diff \cosh(n \theta))}{\diff \theta} = n \sinh(n \theta) \; , \\
\frac{\diff (1 + n^2 (\cosh(\theta) - 1))}{\diff \theta} & = n^2 \sinh(\theta) \; .
\end{align*}
The inequality \eqref{equation:hyperbolic-chebyshev-derivative-inequality} is thus equivalent to
$$
\sinh(n \theta) \ge n \sinh(\theta) \qquad \text{for all $\theta \ge 0$.}
$$
Tho prove this inequality we use the summation formula
$$
\sinh(\alpha + \beta) = \sinh(\alpha) \cosh(\beta) + \sinh(\beta) \cosh(\beta)
$$
and the observation that for $\theta \ge 0$, $\cosh(\theta) \ge 1$ and $\sinh(\theta) \ge 0$. Thus,
$$
\sinh(\alpha + \beta) \ge \sinh(\alpha) + \sinh(\beta) \qquad \text{for $\alpha, \beta \ge 0$.}
$$
Thus by induction on $n$, we see that $\sinh(n \theta) \ge n \sinh(\theta)$ for all $\theta \ge 0$.
